A Technique for Inter-Catalogue Recommendations

### 2014-06-11 paper2/paper.tex
Add popularity histogram and explanation

\subsection*{\em LatentFeatureRecommender}
Although the $NeighborhoodRecommender$ leverages the information contained in
the recommendation graphs to some extent, it is still primarily based on product
descriptions. Presumably, the graphs are derivative of a collaborative filtering
scheme, and hence, reveal associations between products that are not evident
from the language of their descriptions. For example, a dress's description
might not capture if it is more conservative or more risque. Yet, this could be
the salient feature it has in common with an associated recommendation.  These
sort of abstract, perhaps ineffable, features can be essential to item-item
relations. We must therefore ask whether we can more fully utilize the given
graphical information to discover these latent features in order to choose
better inter-catalogue associations.

@@

\subsubsection*{Popularity}
The $LatentFeatureRecommender$ can be significantly improved by incorporating
the notion of item {\em popularity}. The idea is that a more popular item is
more likely to appear as a recommendation independent of how similar it is.  We
see signs of this in the undirected recommendation graphs from www.macys.com.
Although each product has at most four outgoing recommendations on its details
page, we see that some products have over one hundred incoming recommendations.
But how might we define {\em popularity}? A reasonable definition would
introduce a quantity that is correlated with the number of users that indicate a
preference for an item. Like all user data, this is not publicly available
information.  Once again, however, we can leverage the recommendation graph to
infer an approximate popularity metric for each item.

@@

We begin with an attempt to seprate the notion of item similarity from that of
item popularity. The idea that more popular items are more likely to be
recommended than less popular items motivates an alternative mode for conducting
our random walks over a recommendation graph in order to better isolate item
similarity. We do this by adjusting the probability of transitioning to an item
on the graph by making it inversely proportional to that item's popularity
$\rho$. Stationary distributions learned from these alternative random walks
better reflect item-item similarity. We find that removing item popularity from
our simulated user data results in better success when mapping between the
derived concept spaces (as evidenced by the improved precision and recall
discussed in the next section).

@@

>>>>

\subsection*{\em LatentFeatureRecommender}
Although the $NeighborhoodRecommender$ leverages the information contained in
the recommendation graphs to some extent, it is still primarily based on product
descriptions. Presumably, the graphs are derivative of a collaborative filtering
scheme, and hence, reveal associations between products that are not evident
from the language of their descriptions. For example, a dress's description
might not capture whether it is more conservative or more risque. Yet, this
could be the salient feature it has in common with an associated recommendation.
These sort of abstract, perhaps ineffable, features can be essential to
item-item relations. We must therefore ask whether we can more fully utilize the
given graphical information to discover these latent features in order to choose
better inter-catalogue associations.

@@

\subsubsection*{Popularity}
The $LatentFeatureRecommender$ can be significantly improved by incorporating
the notion of item {\em popularity}. The idea is that a more popular item is
more likely to be recommended, independent of how similar it is. We see signs of
this in the undirected recommendation graphs from www.macys.com. Although each
product has at most four outgoing recommendations on its details page, we see
that some products have over one hundred incoming recommendations. But how
might we define {\em popularity}? A reasonable definition would introduce a
quantity that is correlated with the number of users that indicate a preference
for an item. Like all user data, this is not publicly available information.
Once again, however, we can leverage the recommendation graph to infer an
approximate popularity metric for each item.

@@

We begin with an attempt to seprate the notion of item similarity from that of
item popularity. The idea that more popular items are more likely to be
recommended than less popular items motivates an alternative mode for conducting
our random walks over a recommendation graph in order to better isolate item
similarity. We do this by adjusting the probability of transitioning to an item
on the graph by making it inversely proportional to that item's popularity,
$\rho$. Stationary distributions learned from these alternative random walks
better reflect item-item similarity. We find that removing item popularity from
our simulated user data results in better success when mapping between the
derived concept spaces (as evidenced by the improved precision and recall
discussed in the next section).

@@

This assumption is corroborated in the data. The following histogram shows the
number of gold-standard recommendations connected to the top 25 most popular
items, ranked by their value of $\rho$. The histogram also compares the
distribution of predicted recommendations to these items by the
$LatentFeatureRecomender$ with and without popularity taken into account. With
popularity, the predictions follow a similar distribution to the gold-standard
recommendations over these popular items.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=1.0\textwidth]{PopularityHistogram.png}
	\caption{The top 25 most popular items and how many times they appear in
    recommendations. Here we compare the distribution of gold-standard
    recommendations (red) to the recommendations predicted by the
    $LatentFeatureRecommender$ with (blue) and without (yellow) popularity. The
    yellow bars are small and barely visible.}
    \label{fig:PopularityHistogram}
\end{figure}

Note that without popularity, the recommendations are scattered more uniformly
over the thousands of items in the two catalogues, and so the number of times
these 25 items are recommended is significantly smaller (hardly visible in the
plot). This is because the recommendations are soley based on the similarity in
latent concept space and popular items are not inherently more similar to other
items. This recommender in fact still performs better than the 
$ContentBasedRecommender$, however it fails to exploit this trend in the data.
